Quantitative Auditing of AI Fairness with Differentially Private Synthetic Data

### 2025-01-07 paper/paper.tex
Added TODOs

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Auditing AI systems often requires sensitive data, raising security and privacy concerns. To address this, we propose a framework that leverages differentially private synthetic data for fairness evaluations. This approach enables auditors to assess AI systems without exposing sensitive information. Through experiments on real datasets like Adult, COMPAS, and Diabetes, we compare fairness metrics of synthetic and real data. Our results show that synthetic data accurately reflect the fairness properties of real data to a reasonable extent.
\end{abstract}

@@

@@

For the competition, \cite{mckenna2021winning} further manually added certain cliques based on his investigation of the competition dataset. For example, they manually added the clique (sex,city,income). In addition, they would add some edges based on some sophisticated heuristics tailored to that particular dataset. We do not consider these additional heuristics for the generality of our approach. Meanwhile, his other apporach where the auditor does access the real dataset does not fit our framework.

While synthetic data generation methods often leverage domain knowledge to improve accuracy, we chose to employ a vanilla approach with only the maximum spanning tree method without incorporating additional heuristics. This decision was made due to two main factors: first, the need for more specific domain knowledge about the dataset makes it difficult to identify meaningful dependencies; second, prediction targets may differ from dataset to dataset, which means that tailored heuristics could not be effectively generalized. By avoiding dataset-specific adjustments, our method remains general and applicable to a wide range of datasets, in addition to being more computatioanlly efficient.

@@

Next, several AI models are extracted from various real-life authors from Kaggle. They are finetuned to perform well on the original dataset. For one, a random forest model is finetuned by searching hyperparameters settings\cite{Ipbyrne2023}. Another random forest model is finetuned with over-sampling methods\cite{Jawat2024}. Also, a logistic regression model is finetuned by performing principal component analysis\cite{Prashant1112023}.

@@

@@

Although further experiments showed that linking $(S,Y)$ before fitting can somewhat alleviate this problem, the auditor may not always have the background knowledge to identify them; moreover, this improvement is not always guaranteed.

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.

% TODO: add to abstract

\begin{abstract}
Auditing AI systems often requires sensitive data, raising security and privacy concerns. To address this, we propose a framework that leverages differentially private synthetic data for fairness evaluations. This approach enables auditors to assess AI systems without exposing sensitive information. Through experiments on real datasets like Adult, COMPAS, and Diabetes, we compare fairness metrics of synthetic and real data. Our results show that synthetic data accurately reflect the fairness properties of real data to a reasonable extent.
\end{abstract}

@@

% TODO: add to intro to fill 14 pgs

@@

% TODO: reverse order

For the competition, \cite{mckenna2021winning} further manually added certain cliques based on his investigation of the competition dataset. For example, they would add some edges based on some sophisticated heuristics tailored to that particular dataset. In addition, they manually added the clique (sex,city,income) for the prediction target of income. We do not consider these additional ___ for the generality of our approach. Meanwhile, his other apporach where the auditor does not access the real dataset does not fit our framework.

While synthetic data generation methods often leverage domain knowledge to improve accuracy, we chose to employ a vanilla approach with only the maximum spanning tree method without incorporating additional heuristics. This decision was made due to two main factors: first, the specific domain knowledge about the dataset may not be always readily obvious; second, prediction targets may differ from dataset to dataset, which means that tailored heuristics could not be effectively generalized. By avoiding dataset-specific adjustments, our method remains general and applicable to a wide range of datasets, in addition to being more computatioanlly efficient.

@@

Next, several AI models are extracted from various authors from Kaggle. They are finetuned to perform well on the original dataset. For one, a random forest model is finetuned by searching hyperparameters settings\cite{Ipbyrne2023}. Another random forest model is finetuned with over-sampling methods\cite{Jawat2024}. Also, a logistic regression model is finetuned by performing principal component analysis\cite{Prashant1112023}.

% TODO: avg of 100 runs

@@

% TODO: clarify "by common sense"

@@

% Although further experiments showed that linking $(S,Y)$ before fitting can somewhat alleviate this problem. Moreover, this improvement is not always guaranteed.
